---
title: "NSF Collaboration List from PubMed"
author: "Elana J Fertig"
date: "August 15, 2017"
output: word_document
---

# R session

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library('easyPubMed')
library('plyr')
sessionInfo()
```

# Set author for query

This code sets the author from which to query. By default it uses my name, Elana Fertig, in the variable `author`. Please format your name as `"LAST NAME" "FIRST INITIAL"` to match pubmed query structure. 

`authorFilter` contains the full names as `"LAST NAME" "INITALS"` to filter out additional authors with other initials. This may require further customization depending on your name. 

```{r}
author <- 'Fertig E'
authorFilter <- 'Fertig ET'
```

# Establish date range

NSF requires a "list of all persons alphabetically who are current or past collaborators on a project, book, article, rpt, abstract or paper for past 48 mos." By default, we assume that the 48 month period starts on the date of running this software. This can be changed to a different date by storing the desired date in the `currentDate` variable as a `Date` object.

```{r}
currentDate <- Sys.Date()
queryDate <- seq(Sys.Date(), length = 2, by = "-48 months")[2]
```

# Query pubmed

```{r}

pmquery <- sapply(articles_to_list(fetch_pubmed_data(get_pubmed_ids(paste0(author,'[Author]')))),article_to_df)
names(pmquery) <- NULL

# filter out oddly formatted references
pmquery <- pmquery[sapply(pmquery,function(x){any(colnames(x)=='year')})]

```

# Find publications from valid date range

```{r}


# get date of publication
datePMID <- as.Date(sapply(pmquery,function(x){paste(x[1,'year'],
                                               x[1,'month'],
                                              x[1,'day'],sep="-")}))

pmquery <- pmquery[datePMID >= queryDate]

```

# Exclude manuscripts by authors with similar names

Note -- this part of the code may need to be customized depending on your name and additional publications that arise from a pubmed query.

```{r}
getInitials <- function(s) {
  sapply(strsplit(s,split=" "), function(x){
      toupper(paste(substring(x, 1, 1), collapse = ""))
  })
}

validIDs <- !as.logical(sapply(sapply(pmquery,function(x){paste(x$lastname,getInitials(x$firstname))}),function(x){length(grep(authorFilter,x)>0)}))
pmquery <- pmquery[validIDs]
```

# Sort into data frame

```{r}

pmquery.dataframe <- ldply(pmquery,data.frame)

# format author names
pmquery.dataframe$Initials <- getInitials(pmquery.dataframe$firstname)

# find unique co-authors
pmquery.dataframe <- pmquery.dataframe[!duplicated(paste(pmquery.dataframe$lastname,pmquery.dataframe$Initials)),]

# alphabetize the list
pmquery.dataframe <- pmquery.dataframe[order(pmquery.dataframe$lastname,pmquery.dataframe$Initials),]
```

# Word formatted collaboration list

`r paste(paste(paste(pmquery.dataframe$Initials,pmquery.dataframe$lastname),pmquery.dataframe$address,sep=", "),collapse='; ')`
